A fast and memory-efficient transition state search method for complex energy landscapes

### 2022-06-04 manuscript.tex
JCP formatting

\documentclass[twocolumn,10pt]{revtex4}

% Maths
% \usepackage{bm}
\usepackage{amsmath}
\newcommand{\bm}[1]{\boldsymbol{\mathbf{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\grad}{\bm{\nabla}}
\newcommand{\upDelta}{\mathop{}\!\Delta}

% Text
\usepackage[group-separator={,}]{siunitx}
\usepackage{enumitem}
\usepackage{xcolor}

@@

\begin{document}
\title{A fast and memory-efficient transition state search method for complex energy landscapes}
\author{Samuel J. Avis}
\author{Jack R. Panter}
\email[]{j.r.panter@durham.ac.uk}
\author{Halim Kusumaatmaja}
\email[]{halim.kusumaatmaja@durham.ac.uk}
\affiliation{Department of Physics, Durham University, South Road, Durham DH1 3LE, UK}

@@

\section{Results and discussion}
\subsection{BITSS method}
\begin{figure}[tb]
  \includegraphics[width=\linewidth]{toy2d.pdf}
  \caption{\label{fig:toy2d}
    \textbf{Schematics of the BITSS method on a simple 2D potential with two minima.}
    (a) The orange line shows the trajectories of the two states from the minima (blue) to the transition states (red) under the BITSS method.
        The minimum energy pathway is shown by the dashed line.
    (b) A snapshot of the BITSS minimisation showing the driving forces on each state due to the energy constraint, $\bm{F}_\mathrm{E}$, and distance constraint, $\bm{F}_\mathrm{D}$, with $E_1 < E_2$ and $d(\bm{x}_1,\bm{x}_2) < d_i$.
    (c) The final configuration of the BITSS method showing the two states in orange, the transition state in red, and the negative curvature eigenvector, $\bm{\hat{\tau}}$.
  }
\end{figure}

@@

\subsection{Other bracketing methods}
\begin{figure}[htb]
  \includegraphics[width=\linewidth]{bracketing.pdf}
  \caption{\label{fig:bracketing}
    \textbf{Trajectories of the bracketing methods on a hooked potential with a single saddle point.}
    The minimum energy pathway is shown by the dashed line.
  }
\end{figure}
The BITSS potential in \cref{eq:bitss} and iterative steps above offer key advantages over existing bracketing methods that also use two states to locate the transition state.
For instance, in the ridge method \cite{Ionova1993}, the two images are initially chosen to bracket the largest energy point on an interpolated path between the two endpoints.
However, this is not guaranteed to be on the ridge containing the transition state, and specialist methods are required to avoid high-energy local maxima, or when the initial path contains multiple candidate maxima.
In another example, the double ended surface walking method \cite{Zhang2013} requires Gaussian bias potentials to be added at each iteration to force two dimers to climb uphill in the landscape.
For high numbers of degrees of freedom and many iterations, this becomes very computationally expensive.

@@

\subsection{Speed Comparison}
\begin{figure*}[tb]
  \includegraphics[width=\linewidth]{speedtest.pdf}
  \caption{\label{fig:speedtest}
    \textbf{Comparisons between BITSS, DNEB, and the string method in the convergence to the transition state as a function of the number of calculations of the energy gradient.}
    Three test systems are used: (a) a Lennard-Jones seven-particle cluster, (b) cylindrical shell buckling, and (c) wetting of a chemically-striped surface.
    The configurations shown correspond to the two minimum energy states and the transition state, marked by an asterisk.
    The string method and DNEB are repeated with a differing number of images, as listed in the legend in (c).
  }
\end{figure*}

Although reliable, chain-of-states approaches typically use tens of states.
As a result of the BITSS method using only two states, it has the potential to be significantly faster.
This is tested here by measuring the speed of convergence to the transition state for BITSS in comparison with the string method and DNEB for three diverse systems, exhibiting a broad range of energy landscapes.
Details of each are presented in the methods section.

@@

\subsection{Adaptive discretisation}\label{sec:adaptive}
\begin{figure}[tb]
  \includegraphics[width=\linewidth]{adaptivemesh.pdf}
  \caption{\label{fig:adaptivemesh}
    \textbf{Demonstration of BITSS addressing the challenges associated with adaptive remeshing.}
    (a) Snapshots of the BITSS method for the buckling of a cylinder with a changing mesh.
        The radial displacement relative the unbuckled cylinder is shown, as well as the underlying triangular mesh.
    (b) Snapshots for the striped wetting example with different resolutions for the two states.
        Each grid cell denotes 50x50 lattice nodes.
        The zoomed axis shows the difference in the fluid interface between the two final states, as well as the approximated transition state (solid black line).
        This is compared to the transition state found using a high resolution (dashed line).
  }
\end{figure}

@@

\subsection{Complex landscapes}
\begin{figure}[tb]
  \includegraphics[width=\linewidth]{flatdiscontinuous.pdf}
  \caption{\label{fig:flatdiscontinuous}
    \textbf{Demonstration of BITSS applied to flat and discontinuous potentials.}
    (a) Energy profile of the BITSS pathway on a 2D potential with flat regions.
        Blue and red dots denote the minima and transition state, respectively.
        Points of interest are labelled by i--iii (see text).
        Top inset: A zoomed in view around the transition state.
        Bottom inset: The pathway taken, with the edges of the flat regions marked by dashed lines.
    (b) The discontinuous hard-core pair potential used in the seven-particle cluster (orange).
        The standard Lennard-Jones potential is also shown in grey.
    (c) Disconnectivity graphs of the energy landscapes for the seven-particle cluster with the two potentials.
        The two graphs are offset for visibility.
  }
\end{figure}

@@

\subsection{BITSS: Changes for undefined gradients}
A couple of alterations to the method must be made to account for situations where the gradients are unknown.
First, the calculation of $\kappa_d$ in \cref{eq:kd} must be adapted to avoid the use of gradients.
This can be done by simply removing the first term and just using the second term in the equation.
Secondly, L-BFGS can no longer be used because it requires knowledge of the gradients.
We must instead use a minimiser that does not require a differentiable optimisation function, for which we use simulated annealing \cite{Kirkpatrick1983}.
This has a chance of randomly jumping one state over the dividing barrier, but we can reduce this probability by limiting the initial temperature and maximum random displacement.
We typically employ $T_0 = E_\mathrm{B} / 10$, and $d_\mathrm{max} = d(\bm{x}_1, \bm{x}_2) / 100$.

@@

\begin{figure}[htb]
  \includegraphics{fig/BAHschematic.pdf}
  \caption{\label{fig:BAHschematic}
    \textbf{Schematic of the bar and hinge model showing the relevant parameters for a single hinge element.}
    $h_i$ and $\bm{\hat{n}}_i$ denote the height and unit normal of each triangle respectively.
  }
\end{figure}
The gradient of the energy can be obtained by individually considering the stretching and bending energies of a single bond and hinge.
For simplicity, we will ignore the index for the bond and hinge.
Using the variables shown in the schematic in \cref{fig:BAHschematic}, the gradient of the stretching energy of the bond between $\bm{x}_2$ and $\bm{x}_3$ is given by,
\begin{equation}
  \frac{\partial E^\mathrm{S}}{\partial \bm{x}_2} = - \frac{\partial E^\mathrm{S}}{\partial \bm{x}_3} =
    2 k^\mathrm{S} (r - r^0) (\bm{x_2} - \bm{x_3}).
\end{equation}
The gradients of the bending energy of the hinge are
\begin{align}
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_1} &= k^\mathrm{B} \sin(\theta - \theta^0) \frac{\bm{\hat{n}}_a}{h_a}, \\
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_2} &= - k^\mathrm{B} \sin(\theta - \theta^0) \left[\frac{\bm{\hat{n}}_a}{h_a} + \frac{\bm{\hat{n}}_a + \bm{\hat{n}}_b}{r}\right], \\
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_3} &= - k^\mathrm{B} \sin(\theta - \theta^0) \left[\frac{\bm{\hat{n}}_b}{h_b} + \frac{\bm{\hat{n}}_a + \bm{\hat{n}}_b}{r}\right], \\
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_4} &= k^\mathrm{B} \sin(\theta - \theta^0) \frac{\bm{\hat{n}}_b}{h_b}.
\end{align}

@@

@@

\section*{Code availability}
The code used in this study is available from the corresponding author upon reasonable request.

%% start bib
% % \bibliography{../../library.bib}
% \bibliography{bibexport.bib}
% \bibliographystyle{naturemag}
%% mid bib
\begin{thebibliography}{10}
\expandafter\ifx\csname url\endcsname\relax
  \def\url#1{\texttt{#1}}\fi
\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
\providecommand{\bibinfo}[2]{#2}
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem{Bryngelson1995}
\bibinfo{author}{Bryngelson, J.~D.}, \bibinfo{author}{Onuchic, J.~N.},
  \bibinfo{author}{Socci, N.~D.} \& \bibinfo{author}{Wolynes, P.~G.}
\newblock \bibinfo{title}{{Funnels, pathways, and the energy landscape of
  protein folding: A synthesis}}.
\newblock \emph{\bibinfo{journal}{Proteins: Structure, Function, and
  Bioinformatics}} \textbf{\bibinfo{volume}{21}}, \bibinfo{pages}{167--195}
  (\bibinfo{year}{1995}).

\bibitem{Onuchic1997}
\bibinfo{author}{Onuchic, J.~N.}, \bibinfo{author}{Luthey-Schulten, Z.} \&
  \bibinfo{author}{Wolynes, P.~G.}
\newblock \bibinfo{title}{{Theory of Protein Folding: The Energy Landscape
  Perspective}}.
\newblock \emph{\bibinfo{journal}{Annual Review of Physical Chemistry}}
  \textbf{\bibinfo{volume}{48}}, \bibinfo{pages}{545--600}
  (\bibinfo{year}{1997}).

\bibitem{Boehr2006}
\bibinfo{author}{Boehr, D.~D.}, \bibinfo{author}{McElheny, D.},
  \bibinfo{author}{Dyson, H.~J.} \& \bibinfo{author}{Wrightt, P.~E.}
\newblock \bibinfo{title}{{The dynamic energy landscape of dihydrofolate
  reductase catalysis}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{313}},
  \bibinfo{pages}{1638--1642} (\bibinfo{year}{2006}).

\bibitem{Kerns2015}
\bibinfo{author}{Kerns, S.~J.} \emph{et~al.}
\newblock \bibinfo{title}{{The energy landscape of adenylate kinase during
  catalysis}}.
\newblock \emph{\bibinfo{journal}{Nature Structural and Molecular Biology}}
  \textbf{\bibinfo{volume}{22}}, \bibinfo{pages}{124--131}
  (\bibinfo{year}{2015}).

\bibitem{Guo2018a}
\bibinfo{author}{Guo, C.}, \bibinfo{author}{Wang, Z.}, \bibinfo{author}{Wang,
  D.}, \bibinfo{author}{Wang, H.~F.} \& \bibinfo{author}{Hu, P.}
\newblock \bibinfo{title}{{First-Principles Determination of CO Adsorption and
  Desorption on Pt(111) in the Free Energy Landscape}}.
\newblock \emph{\bibinfo{journal}{Journal of Physical Chemistry C}}
  \textbf{\bibinfo{volume}{122}}, \bibinfo{pages}{21478--21483}
  (\bibinfo{year}{2018}).

\bibitem{Richardson2016}
\bibinfo{author}{Richardson, J.~O.} \emph{et~al.}
\newblock \bibinfo{title}{{Concerted hydrogen-bond breaking by quantum
  tunneling in the water hexamer prism}}.
\newblock \emph{\bibinfo{journal}{Science}} \textbf{\bibinfo{volume}{351}},
  \bibinfo{pages}{1310} (\bibinfo{year}{2016}).

\bibitem{Vaillant2019}
\bibinfo{author}{Vaillant, C.~L.}, \bibinfo{author}{Althorpe, S.~C.} \&
  \bibinfo{author}{Wales, D.~J.}
\newblock \bibinfo{title}{{Path Integral Energy Landscapes for Water
  Clusters}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Theory and Computation}}
  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{33--42}
  (\bibinfo{year}{2019}).

\bibitem{Richard2018}
\bibinfo{author}{Richard, D.} \& \bibinfo{author}{Speck, T.}
\newblock \bibinfo{title}{{Crystallization of hard spheres revisited. I.
  Extracting kinetics and free energy landscape from forward flux sampling}}.
\newblock \emph{\bibinfo{journal}{The Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{148}}, \bibinfo{pages}{124110}
  (\bibinfo{year}{2018}).

\bibitem{Wales1998}
\bibinfo{author}{Wales, D.~J.}, \bibinfo{author}{Miller, M.~A.} \&
  \bibinfo{author}{Walsh, T.~R.}
\newblock \bibinfo{title}{{Archetypal energy landscapes}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{394}},
  \bibinfo{pages}{758--760} (\bibinfo{year}{1998}).

\bibitem{Wales2012}
\bibinfo{author}{Wales, D.~J.}
\newblock \bibinfo{title}{{Decoding the energy landscape: Extracting structure,
  dynamics and thermodynamics}}.
\newblock \emph{\bibinfo{journal}{Philosophical Transactions of the Royal
  Society A}} \textbf{\bibinfo{volume}{370}}, \bibinfo{pages}{2877--2899}
  (\bibinfo{year}{2012}).

\bibitem{Panter2019}
\bibinfo{author}{Panter, J.~R.}, \bibinfo{author}{Chen, J.},
  \bibinfo{author}{Zhang, T.} \& \bibinfo{author}{Kusumaatmaja, H.}
\newblock \bibinfo{title}{{Harnessing energy landscape exploration to control
  the buckling of cylindrical shells}}.
\newblock \emph{\bibinfo{journal}{Communications Physics}}
  \textbf{\bibinfo{volume}{2}}, \bibinfo{pages}{151} (\bibinfo{year}{2019}).

\bibitem{Hutchinson2018}
\bibinfo{author}{Hutchinson, J.~W.} \& \bibinfo{author}{Thompson, J. M.~T.}
\newblock \bibinfo{title}{{Imperfections and energy barriers in shell
  buckling}}.
\newblock \emph{\bibinfo{journal}{International Journal of Solids and
  Structures}} \textbf{\bibinfo{volume}{148-149}}, \bibinfo{pages}{157--168}
  (\bibinfo{year}{2018}).

\bibitem{Zhang2014}
\bibinfo{author}{Zhang, Y.} \& \bibinfo{author}{Ren, W.}
\newblock \bibinfo{title}{{Numerical study of the effects of surface topography
  and chemistry on the wetting transition using the string method}}.
\newblock \emph{\bibinfo{journal}{The Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{141}}, \bibinfo{pages}{244705}
  (\bibinfo{year}{2014}).

\bibitem{Panter2019b}
\bibinfo{author}{Panter, J.~R.}, \bibinfo{author}{Gizaw, Y.} \&
  \bibinfo{author}{Kusumaatmaja, H.}
\newblock \bibinfo{title}{{Multifaceted design optimization for superomniphobic
  surfaces}}.
\newblock \emph{\bibinfo{journal}{Science Advances}}
  \textbf{\bibinfo{volume}{5}}, \bibinfo{pages}{eaav7328}
  (\bibinfo{year}{2019}).

\bibitem{Amabili2017}
\bibinfo{author}{Amabili, M.}, \bibinfo{author}{Giacomello, A.},
  \bibinfo{author}{Meloni, S.} \& \bibinfo{author}{Casciola, C.~M.}
\newblock \bibinfo{title}{{Collapse of superhydrophobicity on nanopillared
  surfaces}}.
\newblock \emph{\bibinfo{journal}{Physical Review Fluids}}
  \textbf{\bibinfo{volume}{2}}, \bibinfo{pages}{034202} (\bibinfo{year}{2017}).

\bibitem{Othayoth2020}
\bibinfo{author}{Othayoth, R.}, \bibinfo{author}{Thoms, G.} \&
  \bibinfo{author}{Li, C.}
\newblock \bibinfo{title}{{An energy landscape approach to locomotor
  transitions in complex 3D terrain}}.
\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
  Sciences of the United States of America}} \textbf{\bibinfo{volume}{117}},
  \bibinfo{pages}{14987--14995} (\bibinfo{year}{2020}).

\bibitem{Filipov2015}
\bibinfo{author}{Filipov, E.~T.}, \bibinfo{author}{Tachi, T.},
  \bibinfo{author}{Paulino, G.~H.} \& \bibinfo{author}{Weitz, D.~A.}
\newblock \bibinfo{title}{{Origami tubes assembled into stiff, yet
  reconfigurable structures and metamaterials}}.
\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
  Sciences of the United States of America}} \textbf{\bibinfo{volume}{112}},
  \bibinfo{pages}{12321} (\bibinfo{year}{2015}).

\bibitem{Zhai2018}
\bibinfo{author}{Zhai, Z.}, \bibinfo{author}{Wang, Y.} \&
  \bibinfo{author}{Jiang, H.}
\newblock \bibinfo{title}{{Origami-inspired, on-demand deployable and
  collapsible mechanical metamaterials with tunable stiffness}}.
\newblock \emph{\bibinfo{journal}{Proceedings of the National Academy of
  Sciences of the United States of America}} \textbf{\bibinfo{volume}{115}},
  \bibinfo{pages}{2032--2037} (\bibinfo{year}{2018}).

\bibitem{Bertoldi2017}
\bibinfo{author}{Bertoldi, K.}, \bibinfo{author}{Vitelli, V.},
  \bibinfo{author}{Christensen, J.} \& \bibinfo{author}{{Van Hecke}, M.}
\newblock \bibinfo{title}{{Flexible mechanical metamaterials}}.
\newblock \emph{\bibinfo{journal}{Nature Reviews Materials}}
  \textbf{\bibinfo{volume}{2}}, \bibinfo{pages}{17066} (\bibinfo{year}{2017}).

\bibitem{Truby2016}
\bibinfo{author}{Truby, R.~L.} \& \bibinfo{author}{Lewis, J.~A.}
\newblock \bibinfo{title}{{Printing soft matter in three dimensions}}.
\newblock \emph{\bibinfo{journal}{Nature}} \textbf{\bibinfo{volume}{540}},
  \bibinfo{pages}{371--378} (\bibinfo{year}{2016}).

\bibitem{Chi2022}
\bibinfo{author}{Chi, Y.} \emph{et~al.}
\newblock \bibinfo{title}{{Bistable and Multistable Actuators for Soft Robots:
  Structures, Materials, and Functionalities}}.
\newblock \emph{\bibinfo{journal}{Advanced Materials}}
  \textbf{\bibinfo{volume}{34}}, \bibinfo{pages}{2110384}
  (\bibinfo{year}{2022}).

\bibitem{Bonfanti2020}
\bibinfo{author}{Bonfanti, S.}, \bibinfo{author}{Guerra, R.},
  \bibinfo{author}{Font-Clos, F.}, \bibinfo{author}{Rayneau-Kirkhope, D.} \&
  \bibinfo{author}{Zapperi, S.}
\newblock \bibinfo{title}{{Automatic design of mechanical metamaterial
  actuators}}.
\newblock \emph{\bibinfo{journal}{Nature Communications}}
  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{4162} (\bibinfo{year}{2020}).

\bibitem{Shan2015}
\bibinfo{author}{Shan, S.} \emph{et~al.}
\newblock \bibinfo{title}{{Multistable Architected Materials for Trapping
  Elastic Strain Energy}}.
\newblock \emph{\bibinfo{journal}{Advanced Materials}}
  \textbf{\bibinfo{volume}{27}}, \bibinfo{pages}{4296--4301}
  (\bibinfo{year}{2015}).

\bibitem{Giri2021}
\bibinfo{author}{Giri, T.~R.} \& \bibinfo{author}{Mailen, R.}
\newblock \bibinfo{title}{{Controlled snapping sequence and energy absorption
  in multistable mechanical metamaterial cylinders}}.
\newblock \emph{\bibinfo{journal}{International Journal of Mechanical
  Sciences}} \textbf{\bibinfo{volume}{204}}, \bibinfo{pages}{106541}
  (\bibinfo{year}{2021}).

\bibitem{Cerjan1981}
\bibinfo{author}{Cerjan, C.~J.} \& \bibinfo{author}{Miller, W.~H.}
\newblock \bibinfo{title}{{On finding transition states}}.
\newblock \emph{\bibinfo{journal}{The Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{75}}, \bibinfo{pages}{2800--2806}
  (\bibinfo{year}{1981}).

\bibitem{Heyden2005}
\bibinfo{author}{Heyden, A.}, \bibinfo{author}{Bell, A.~T.} \&
  \bibinfo{author}{Keil, F.~J.}
\newblock \bibinfo{title}{{Efficient methods for finding transition states in
  chemical reactions: Comparison of improved dimer method and partitioned
  rational function optimization method}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{123}}, \bibinfo{pages}{224101}
  (\bibinfo{year}{2005}).

\bibitem{Kastner2008}
\bibinfo{author}{K{\"{a}}stner, J.} \& \bibinfo{author}{Sherwood, P.}
\newblock \bibinfo{title}{{Superlinearly converging dimer method for transition
  state search}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{128}}, \bibinfo{pages}{014106}
  (\bibinfo{year}{2008}).

\bibitem{Zhang2016}
\bibinfo{author}{Zhang, L.}, \bibinfo{author}{Du, Q.} \&
  \bibinfo{author}{Zheng, Z.}
\newblock \bibinfo{title}{{Optimization-based Shrinking Dimer Method for
  Finding Transition States}}.
\newblock \emph{\bibinfo{journal}{SIAM Journal on Scientific Computing}}
  \textbf{\bibinfo{volume}{38}}, \bibinfo{pages}{A528--A544}
  (\bibinfo{year}{2016}).

\bibitem{E2007}
\bibinfo{author}{E, W.}, \bibinfo{author}{Ren, W.} \&
  \bibinfo{author}{Vanden-Eijnden, E.}
\newblock \bibinfo{title}{{Simplified and improved string method for computing
  the minimum energy paths in barrier-crossing events}}.
\newblock \emph{\bibinfo{journal}{The Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{126}}, \bibinfo{pages}{164103}
  (\bibinfo{year}{2007}).

\bibitem{Ren2013}
\bibinfo{author}{Ren, W.} \& \bibinfo{author}{Vanden-Eijnden, E.}
\newblock \bibinfo{title}{{A climbing string method for saddle point search}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{138}}, \bibinfo{pages}{134105}
  (\bibinfo{year}{2013}).

\bibitem{E2002}
\bibinfo{author}{E, W.}, \bibinfo{author}{Ren, W.} \&
  \bibinfo{author}{Vanden-Eijnden, E.}
\newblock \bibinfo{title}{{String method for the study of rare events}}.
\newblock \emph{\bibinfo{journal}{Physical Review B}}
  \textbf{\bibinfo{volume}{66}}, \bibinfo{pages}{052301}
  (\bibinfo{year}{2002}).

\bibitem{Trygubenko2004}
\bibinfo{author}{Trygubenko, S.~A.} \& \bibinfo{author}{Wales, D.~J.}
\newblock \bibinfo{title}{{A doubly nudged elastic band method for finding
  transition states}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{120}}, \bibinfo{pages}{2082--2094}
  (\bibinfo{year}{2004}).

\bibitem{Wales2012a}
\bibinfo{author}{Wales, D.~J.} \& \bibinfo{author}{Carr, J.~M.}
\newblock \bibinfo{title}{{Quasi-Continuous Interpolation Scheme for Pathways
  between Distant Configurations}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Theory and Computation}}
  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{5020--5034}
  (\bibinfo{year}{2012}).

\bibitem{Dewar1984}
\bibinfo{author}{Dewar, M. J.~S.}, \bibinfo{author}{Healy, E.~F.} \&
  \bibinfo{author}{Stewart, J. J.~P.}
\newblock \bibinfo{title}{{Location of transition states in reaction
  mechanisms}}.
\newblock \emph{\bibinfo{journal}{Journal of the Chemical Society, Faraday
  Transactions 2}} \textbf{\bibinfo{volume}{80}}, \bibinfo{pages}{227}
  (\bibinfo{year}{1984}).

\bibitem{Ionova1993}
\bibinfo{author}{Ionova, I.~V.} \& \bibinfo{author}{Carter, E.~A.}
\newblock \bibinfo{title}{{Ridge method for finding saddle points on potential
  energy surfaces}}.
\newblock \emph{\bibinfo{journal}{The Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{98}}, \bibinfo{pages}{6377--6386}
  (\bibinfo{year}{1993}).

\bibitem{Miron2001}
\bibinfo{author}{Miron, R.~A.} \& \bibinfo{author}{Fichthorn, K.~A.}
\newblock \bibinfo{title}{{The Step and Slide method for finding saddle points
  on multidimensional potential surfaces}}.
\newblock \emph{\bibinfo{journal}{The Journal of Chemical Physics}}
  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{8742--8747}
  (\bibinfo{year}{2001}).

\bibitem{Zhang2013}
\bibinfo{author}{Zhang, X.-J.}, \bibinfo{author}{Shang, C.} \&
  \bibinfo{author}{Liu, Z.-P.}
\newblock \bibinfo{title}{{Double-Ended Surface Walking Method for Pathway
  Building and Transition State Location of Complex Reactions}}.
\newblock \emph{\bibinfo{journal}{Journal of Chemical Theory and Computation}}
  \textbf{\bibinfo{volume}{9}}, \bibinfo{pages}{5745--5753}
  (\bibinfo{year}{2013}).

>>>>

\documentclass[aip,jcp,reprint,twocolumn]{revtex4-1}

% Maths
\usepackage{bm}
\usepackage{amsmath}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\grad}{\bm{\nabla}}
\newcommand{\upDelta}{\mathop{}\!\Delta}

% Text
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[group-separator={,}]{siunitx}
\usepackage{enumitem}
\usepackage{xcolor}

@@

% \usepackage[mathlines]{lineno}
% \linenumbers\relax

\begin{document}
\title{A fast and memory-efficient transition state search method for complex energy landscapes}

\author{Samuel J. Avis}
\author{Jack R. Panter}
\email[]{j.r.panter@durham.ac.uk}
\author{Halim Kusumaatmaja}
\email[]{halim.kusumaatmaja@durham.ac.uk}
\affiliation{Department of Physics, Durham University, South Road, Durham DH1 3LE, UK}

\date{\today}

@@

\section{BITSS method}
\begin{figure}[tb]
  \includegraphics[width=\linewidth]{toy2d.pdf}
  \caption{\label{fig:toy2d}
    Schematics of the BITSS method on a simple 2D potential with two minima.
    (a) The orange line shows the trajectories of the two states from the minima (blue) to the transition states (red) under the BITSS method.
        The minimum energy pathway is shown by the dashed line.
    (b) A snapshot of the BITSS minimisation showing the driving forces on each state due to the energy constraint, $\bm{F}_\mathrm{E}$, and distance constraint, $\bm{F}_\mathrm{D}$, with $E_1 < E_2$ and $d(\bm{x}_1,\bm{x}_2) < d_i$.
    (c) The final configuration of the BITSS method showing the two states in orange, the transition state in red, and the negative curvature eigenvector, $\bm{\hat{\tau}}$.
  }
\end{figure}

@@

\subsection{Changes for undefined gradients}
A couple of alterations to the method must be made to account for situations where the gradients are unknown.
First, the calculation of $\kappa_d$ in \cref{eq:kd} must be adapted to avoid the use of gradients.
This can be done by simply removing the first term and just using the second term in the equation.
Secondly, L-BFGS can no longer be used because it requires knowledge of the gradients.
We must instead use a minimiser that does not require a differentiable optimisation function, for which we use simulated annealing \cite{Kirkpatrick1983}.
This has a chance of randomly jumping one state over the dividing barrier, but we can reduce this probability by limiting the initial temperature and maximum random displacement.
We typically employ $T_0 = E_\mathrm{B} / 10$, and $d_\mathrm{max} = d(\bm{x}_1, \bm{x}_2) / 100$.

\section{Results and Discussion}
\subsection{Comparison with other bracketing methods}
\begin{figure}[htb]
  \includegraphics[width=\linewidth]{bracketing.pdf}
  \caption{\label{fig:bracketing}
    Trajectories of the bracketing methods on a hooked potential with a single saddle point.
    The minimum energy pathway is shown by the dashed line.
  }
\end{figure}
The BITSS potential in \cref{eq:bitss} and iterative steps above offer key advantages over existing bracketing methods that also use two states to locate the transition state.
For instance, in the ridge method \cite{Ionova1993}, the two images are initially chosen to bracket the largest energy point on an interpolated path between the two endpoints.
However, this is not guaranteed to be on the ridge containing the transition state, and specialist methods are required to avoid high-energy local maxima, or when the initial path contains multiple candidate maxima.
In another example, the double ended surface walking method \cite{Zhang2013} requires Gaussian bias potentials to be added at each iteration to force two dimers to climb uphill in the landscape.
For high numbers of degrees of freedom and many iterations, this becomes very computationally expensive.

@@

\subsection{Speed comparison}
\begin{figure*}[tb]
  \includegraphics[width=\linewidth]{speedtest.pdf}
  \caption{\label{fig:speedtest}
    Comparisons between BITSS, DNEB, and the string method in the convergence to the transition state as a function of the number of calculations of the energy gradient.
    Three test systems are used: (a) a Lennard-Jones seven-particle cluster, (b) cylindrical shell buckling, and (c) wetting of a chemically-striped surface.
    The configurations shown correspond to the two minimum energy states and the transition state, marked by an asterisk.
    The string method and DNEB are repeated with a differing number of images, as listed in the legend in (c).
  }
\end{figure*}

Although reliable, chain-of-states approaches typically use tens of states.
As a result of the BITSS method using only two states, it has the potential to be significantly faster.
This is tested here by measuring the speed of convergence to the transition state for BITSS in comparison with the string method \cite{E2007} and DNEB \cite{Trygubenko2004} for three diverse systems, exhibiting a broad range of energy landscapes.
Details of each are presented in the methods section.

@@

\subsection{Adaptive discretisation}\label{sec:adaptive}
\begin{figure}[tb]
  \includegraphics[width=\linewidth]{adaptivemesh.pdf}
  \caption{\label{fig:adaptivemesh}
    Demonstration of BITSS addressing the challenges associated with adaptive remeshing.
    (a) Snapshots of the BITSS method for the buckling of a cylinder with a changing mesh.
        The radial displacement relative the unbuckled cylinder is shown, as well as the underlying triangular mesh.
    (b) Snapshots for the striped wetting example with different resolutions for the two states.
        Each grid cell denotes 50x50 lattice nodes.
        The zoomed axis shows the difference in the fluid interface between the two final states, as well as the approximated transition state (solid black line).
        This is compared to the transition state found using a high resolution (dashed line).
  }
\end{figure}

@@

\subsection{Complex landscapes}
\begin{figure}[tb]
  \includegraphics[width=\linewidth]{flatdiscontinuous.pdf}
  \caption{\label{fig:flatdiscontinuous}
    Demonstration of BITSS applied to flat and discontinuous potentials.
    (a) Energy profile of the BITSS pathway on a 2D potential with flat regions.
        Blue and red dots denote the minima and transition state, respectively.
        Points of interest are labelled by i--iii (see text).
        Top inset: A zoomed in view around the transition state.
        Bottom inset: The pathway taken, with the edges of the flat regions marked by dashed lines.
    (b) The discontinuous hard-core pair potential used in the seven-particle cluster (orange).
        The standard Lennard-Jones potential is also shown in grey.
    (c) Disconnectivity graphs of the energy landscapes for the seven-particle cluster with the two potentials.
        The two graphs are offset for visibility.
  }
\end{figure}

@@

@@

\begin{figure}[htb]
  \includegraphics{fig/BAHschematic.pdf}
  \caption{\label{fig:BAHschematic}
    Schematic of the bar and hinge model showing the relevant parameters for a single hinge element.
    $h_i$ and $\bm{\hat{n}}_i$ denote the height and unit normal of each triangle respectively.
  }
\end{figure}
The gradient of the energy can be obtained by individually considering the stretching and bending energies of a single bond and hinge.
For simplicity, we will ignore the index for the bond and hinge.
Using the variables shown in the schematic in \cref{fig:BAHschematic}, the gradient of the stretching energy of the bond between $\bm{x}_2$ and $\bm{x}_3$ is given by,
\begin{equation}
  \frac{\partial E^\mathrm{S}}{\partial \bm{x}_2} = - \frac{\partial E^\mathrm{S}}{\partial \bm{x}_3} =
    2 k^\mathrm{S} (r - r^0) (\bm{x_2} - \bm{x_3}).
\end{equation}
The gradients of the bending energy of the hinge are
\begin{align}
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_1} &= k^\mathrm{B} \sin(\theta - \theta^0) \frac{\bm{\hat{n}}_a}{h_a}, \\
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_2} &= - k^\mathrm{B} \sin(\theta - \theta^0) \left[\frac{\bm{\hat{n}}_a}{h_a} + \frac{\bm{\hat{n}}_a + \bm{\hat{n}}_b}{r}\right], \\
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_3} &= - k^\mathrm{B} \sin(\theta - \theta^0) \left[\frac{\bm{\hat{n}}_b}{h_b} + \frac{\bm{\hat{n}}_a + \bm{\hat{n}}_b}{r}\right], \\
  \frac{\partial E^\mathrm{B}}{\partial \bm{x}_4} &= k^\mathrm{B} \sin(\theta - \theta^0) \frac{\bm{\hat{n}}_b}{h_b}.
\end{align}

@@

\section*{Supplementary material}
See the supplementary material for derivations of the expressions of the constraint coefficients in \cref{eq:kd,eq:ke}, and the demonstration of the BITSS method applied to a path with multiple transition states.

\section*{Conflict of Interest}
The authors have no conflicts to disclose.

\section*{Author Contributions}
All authors conceived and planned the project.
S.~J.~A. performed the numerical experiments and analysed the data.
J.~R.~P. and H.~K. supervised the study.
S.~J.~A. wrote the paper, with review and input from and J.~R.~P. and H.~K.

@@

%% start bib
\bibliographystyle{aipnum4-1}
\bibliography{bibexport.bib}
% % \bibliography{../../library.bib}
%% mid bib
%% end bib
